Enabling GPU Virtualization in Cloud Environments

### 2016-02-02 paper.tex
reviews

@@

Unfortunately, sharing GPU resources among multiple VMs in cloud environments 
is as quite more complex than in physical servers, resulting in a low utilization rate.
On one hand, instances in public clouds are not easily customizable. 
On the other, in a private cloud, although the instances can be customized in many aspects, when referring to GPUs
the number of options is reduced.
Moreover, neither vendors nor tools offer GPGPU services, missing the opportunity of providing flexibility and boost the utilization of GPU resources in both private or public clouds.

@@

From the HPC perspective, OpenStack offers high performance configurations with different hardware architectures. Despite in OpenStack it is possible to 
request special resources, such as GPUs, choosing a host based on the existence
of a GPU is currently unsupported~\cite{OpenStackGPU}.

@@

The work in~\cite{younge2013enabling} allows the VM managed
by the Xen hypervisor to access the GPUs in a physical
node, but with this solution a node cannot 
use more GPUs than those hosted in the machine, and an idle GPU can not be shared with other machines. The solution presented 
by gVirtus~\cite{gvirtus} virtualizes GPUs and makes
them accessible for any VM in the cluster. However, this kind
of virtualization strongly depends on the hypervisor, and so
does its performance. Similar solution is presented in gCloud~\cite{diab2013dynamic}. 
While this solution is not yet integrated in a Cloud Computing Manager, its main
drawback is that the application's code must be modified
in order to run in the virtual-GPU environment. A runtime component to provide abstraction
and sharing of GPUs is presented in~\cite{becchi2012virtual}, which allows scheduling policies 
to isolate and share GPUs in a cluster for a set of applications. 
The work introduced in~\cite{jungpgpu} is more mature; however,
it is only focused on compute-intensive HPC applications.

Our proposal goes further and, in addition to 
bringing solutions for all kind of HPC applications, aims 
to boost flexibility in the use of GPUs.

@@

In both cases, the power of current cloud facilities relies on the use 
of VMs which  
emulate a physical node with its own 
CPUs, storage, network interconnection, GPU and OS. 
However, providing a VM equipped with a GPU implies
that the guest host must be equipped with at least one real GPU, and
once it has been assigned to an instance, no other VM can use
it. Therefore, the number of VMs that can access the GPUs is 
limited by the number of the installed devices. This lack of 
flexibility is also present in cluster configurations, where 
depending on the users' necessities, 
two approaches to enable GPGPU computation can be identified:
the conservative option integrates many hosts with GPUs to 
ensure that it is likely to satisfy all requests; in a
more daring approach, a reduced count of some GPU-enabled nodes is configured, in the hope that a large burst of GPU requests 
 does not exhaust the cluster's resources.

@@

\begin{table}[htb]
\renewcommand{\arraystretch}{1.3}
\caption{Shown HPC instances available in US EAST (N. VIRGINIA) in June 2015.}
\label{table:awsInstances}
\tabcolsep=0.09cm
\begin{center}\begin{tabular}{cccccc}
Name & vCPUs & Memory & Network & GPUs & Price\\ \hline \hline
c3.2xlarge & 8 & 15 GiB & High & 0 & \$ 0.21\\ \hline
c3.8xlarge & 32 & 60 GiB & 10 Gb & 0 & \$ 1.68 \\ \hline
g2.2xlarge & 8 & 15 GiB & High & 1 & \$ 0.65\\ \hline
g2.8xlarge & 32 & 60 GiB & 10 Gb & 4 & \$ 2.6 \\ \hline
\end{tabular}\end{center}\end{table}

@@

\subsubsection{GPUs}
An instance relies on a VM that runs on a real node with its own virtualized components. 
Therefore AWS can leverage a virtualization framework to offer GPU services to all the instances.
Although the {\tt nvidia-smi} command indicates that the GPUs installed are NVIDIA GRID K520, 
we need to verify that these are non-virtualized devices. For this purpose, we execute the NVIDIA SDK {\tt bandwidthtest}. 
As shown in the Table~\ref{table:bwt}, the bandwidth achieved in this test is higher than the network bandwidth, revealing that the accelerator is an actual GPU. 
% \begin{table}[htb]
% \renewcommand{\arraystretch}{1.3}
% \caption{NVIDIA SDK {\tt bandwidthtest} execution transferring 32MB using pageable memory using local GPU.}
% \label{table:bwt}
% \tabcolsep=0.09cm
% \begin{center}\begin{tabular}{cccc}
% Name &  Data Movement & Network & Bandwidth \\ \hline \hline
% g2.2xlarge & Host to Device & High& 3,004 MB/s \\ \hline
% g2.2xlarge & Device to Host & High& 2,809 MB/s\\ \hline
% g2.8xlarge & Host to Device & 10 Gb& 2,814 MB/s\\ \hline
% g2.8xlarge & Device to Host & 10 Gb& 3,182 MB/s\\ \hline
% \end{tabular}\end{center}\end{table}

\begin{table}[htb]
\renewcommand{\arraystretch}{1.3}
\caption{Execution of NVIDIA SDK {\tt bandwidthtest} to transferring 32MB using pageable memory in a local GPU.}
\label{table:bwt}
\tabcolsep=0.09cm
\begin{center}\begin{tabular}{ccc}
Name &  Data Movement & Bandwidth \\ \hline \hline
g2.2xlarge & Host to Device & 3,004 MB/s \\ \hline
g2.2xlarge & Device to Host & 2,809 MB/s\\ \hline
g2.8xlarge & Host to Device & 2,814 MB/s\\ \hline
g2.8xlarge & Device to Host & 3,182 MB/s\\ \hline
\end{tabular}\end{center}\end{table}

\subsection{Testbed Scenarios}
In this section, we introduce the different client-server configurations. 
All scenarios are based on the maximum number of instances 
that a user can freely select without submitting a formal request. 
In particular, the maximum number for ``g2.2xlarge'' is 5; for ``g2.8xlarge'' it is 2.

@@

\begin{table}[htb]
\renewcommand{\arraystretch}{1.3}
\caption{Execution of NVIDIA SDK {\tt bandwidthtest} to transferring 32MB using pageable memory in a remote GPU using {rCUDA}.}
\label{table:bwtrcuda}
\tabcolsep=0.09cm
\begin{center}\begin{tabular}{cccc}
Scenario &  Data Movement & Network & Bandwidth \\ \hline \hline
A & Host-to-Device & High& 127 MB/s \\ \hline
A & Device-to-Host & High& 126 MB/s\\ \hline
B & Host-to-Device & 10 Gb& 858 MB/s\\ \hline
B & Device-to-Host & 10 Gb& 843 MB/s\\ \hline
\end{tabular}\end{center}\end{table}

@@

\subsection{Discussion}
The previous experiments reveal that the AWS GPU-instances are not appropriate for HPC
because neither the network nor the accelerators are powerful enough to deliver high performance
when running compute-intensive parallel applications. 
In other words, AWS is more oriented toward offering a general-purpose service than to provide high performance.
Also, AWS fails in offering flexibility as it forces the user to choose between a basic instance with a GPU and a powerful instance with 4 GPUs.

@@

In terms of GPU-shareability, AWS reserves GPU-capable physical machines which will be idle until a GPU-instance request is made.
Maintaining an expensive accelerator in a standby state is counter-productive. 
It makes more sense to dedicate less devices, accessible from any machine.

>>>>

%\author{Ommitted for blind review}

@@

Unfortunately, sharing GPU resources among multiple VMs in cloud environments is more complex than in physical servers.
On one hand, instances in public clouds are not easily customizable. 
On the other, in a private cloud, although the instances can be customized in many aspects, when referring to GPUs
the number of options is reduced.
As a result, neither vendors nor tools offer GPGPU services.

@@

From the HPC perspective, OpenStack offers high performance configurations with different hardware architectures. 
Even though in OpenStack it is possible to work with GPUs, the Nova project is not currently aware of them~\cite{OpenStackGPU}.

@@

The work in~\cite{younge2013enabling} allows the VM managed
by the Xen hypervisor to access the GPUs in a physical
node, but with this solution a node cannot 
use more GPUs than those hosted in itself, and an idle GPU can not be shared with other nodes. 
The solution presented by gVirtus~\cite{gvirtus} virtualizes GPUs and makes
them accessible for any VM in the cluster. However, this kind
of virtualization strongly depends on the hypervisor, and so
does its performance. Similar solution is presented in gCloud~\cite{diab2013dynamic}. 
While this solution is not yet integrated in a Cloud Computing Manager, its main
drawback is that the application's code must be modified
in order to run in the virtual-GPU environment. A runtime component to provide abstraction
and sharing of GPUs is presented in~\cite{becchi2012virtual}, which allows scheduling policies 
to isolate and share GPUs in a cluster for a set of applications. 
The work introduced in~\cite{jungpgpu} is more mature; however,
it is only focused on compute-intensive HPC applications.

Our proposal goes further, not only bringing solutions for all kind of HPC applications, 
but also aiming to boost flexibility in the use of GPUs.

@@

In both cases, the power of current cloud facilities relies on the use 
of VMs which  
emulate a physical node with its own 
CPUs, storage, network interconnection, GPU and OS. 
However, providing a VM equipped with a GPU implies
that the guest host must be equipped with at least one real GPU, and
once it has been assigned to an instance, no other VM can use
it. Therefore, the number of VMs that can access the GPUs is 
limited by the number of the installed devices. 
This lack of 
flexibility is also present in cluster configurations, where 
depending on the users' necessities, 
two approaches to enable GPGPU computation can be identified.
On the one hand, the conservative option, that integrates many hosts with GPUs to ensure that it is likely to satisfy all requests.
On the other hand, a more daring approach, a reduced count of GPU-enabled nodes is configured, based on an euristhic value calculated from previous experience.

@@

\begin{table}[htb]
\renewcommand{\arraystretch}{1.3}
\caption{Shown HPC instances available in US EAST (N. VIRGINIA) in June 2015.}
\label{table:awsInstances}
\tabcolsep=0.09cm
\begin{center}\begin{tabular}{cccccc}
Name & vCPUs & Memory & Network & GPUs & Price\\ \hline \hline
c3.2xlarge & 8 & 15 GiB & High & 0 & \$ 0.42\\ \hline
c3.8xlarge & 32 & 60 GiB & 10 Gb & 0 & \$ 1.68 \\ \hline
g2.2xlarge & 8 & 15 GiB & High & 1 & \$ 0.65\\ \hline
g2.8xlarge & 32 & 60 GiB & 10 Gb & 4 & \$ 2.6 \\ \hline
\end{tabular}\end{center}\end{table}

@@

\subsubsection{GPUs}
An instance relies on a VM that runs on a real node with its own virtualized components. 
Therefore AWS can leverage a virtualization framework to offer GPU services to all the instances.
Although the {\tt nvidia-smi} command indicates that the GPUs installed are NVIDIA GRID K520, 
we need to verify that these are non-virtualized devices. For this purpose, we execute the NVIDIA SDK {\tt bandwidthtest}. 
As shown in the Table~\ref{table:bwt}, the bandwidth achieved in this test is higher than the network bandwidth, which suggests that the accelerator is an actual GPU. 
% \begin{table}[htb]
% \renewcommand{\arraystretch}{1.3}
% \caption{NVIDIA SDK {\tt bandwidthtest} execution transferring 32MB using pageable memory using local GPU.}
% \label{table:bwt}
% \tabcolsep=0.09cm
% \begin{center}\begin{tabular}{cccc}
% Name &  Data Movement & Network & Bandwidth \\ \hline \hline
% g2.2xlarge & Host to Device & High& 3,004 MB/s \\ \hline
% g2.2xlarge & Device to Host & High& 2,809 MB/s\\ \hline
% g2.8xlarge & Host to Device & 10 Gb& 2,814 MB/s\\ \hline
% g2.8xlarge & Device to Host & 10 Gb& 3,182 MB/s\\ \hline
% \end{tabular}\end{center}\end{table}

\begin{table}[htb]
\renewcommand{\arraystretch}{1.3}
\caption{Results of {\tt bandwidthtest} transferring 32MB using pageable memory in a local GPU.}
\label{table:bwt}
\tabcolsep=0.09cm
\begin{center}\begin{tabular}{ccc}
Name &  Data Movement & Bandwidth \\ \hline \hline
g2.2xlarge & Host to Device & 3,004 MB/s \\ \hline
g2.2xlarge & Device to Host & 2,809 MB/s\\ \hline
g2.8xlarge & Host to Device & 2,814 MB/s\\ \hline
g2.8xlarge & Device to Host & 3,182 MB/s\\ \hline
\end{tabular}\end{center}\end{table}

\subsection{Testbed Scenarios}
In this section, we introduce different client-server configurations. 
All scenarios are based on the maximum number of instances 
that a user can freely select without submitting a formal request. 
In particular, the maximum number for ``g2.2xlarge'' is 5; for ``g2.8xlarge'' it is 2.

@@

\begin{table}[htb]
\renewcommand{\arraystretch}{1.3}
\caption{Results of {\tt bandwidthtest} transferring 32MB using pageable memory in a remote GPU using {rCUDA}.}
\label{table:bwtrcuda}
\tabcolsep=0.09cm
\begin{center}\begin{tabular}{cccc}
Scenario &  Data Movement & Network & Bandwidth \\ \hline \hline
A & Host-to-Device & High& 127 MB/s \\ \hline
A & Device-to-Host & High& 126 MB/s\\ \hline
B & Host-to-Device & 10 Gb& 858 MB/s\\ \hline
B & Device-to-Host & 10 Gb& 843 MB/s\\ \hline
\end{tabular}\end{center}\end{table}

@@

\subsection{Discussion}
The previous experiments reveal that the AWS GPU-instances are not appropriate for HPC
because neither the network nor the accelerators are powerful enough to deliver high performance
when running compute-intensive parallel applications. As~\cite{tonithesis} demonstrates, network and device types are critical in the performance.
In other words, AWS is more oriented toward offering a general-purpose service than to provide high performance.
Also, AWS fails in offering flexibility as it forces the user to choose between a basic instance with a GPU and a powerful instance with 4 GPUs.

@@

In terms of GPU-shareability, AWS reserves GPU-capable physical machines which will be waiting for a GPU-instance request.
Investing in expensive accelerators to keep them in a standby state is counter-productive. 
It makes more sense to dedicate less devices, accessible from any machine, with a higher utilization rate.
